Modeling memory and taxis to analyze migrations

### 2021-01-11 code/StrategicVsTactical_PartI_v2.Rmd
name
delete
Improved model structure.

---
title: "Modeling memory and taxis to analyze migrations"
author: "E. Gurarie, B. Fagan, others"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
theme: cerulean
editor_options:
  chunk_output_type: console
---

@@

```{r, message=FALSE, warning=FALSE}
require(deSolve)
require(ReacTran)
require(magrittr)
source("functions.R")
```

@@

```{r}
ForagingMemoryModel <- function(t, pop, parms, pop_lag, resource, dx){
  tran.1D(C = pop, D = parms["epsilon"], 
                   flux.up = 0, flux.down = 0, 
                   v = parms["alpha"] * diff(resource)/dx + 
                       parms["beta"] * diff(pop_lag)/dx, 
                   dx = dx)
}
```

@@

parameters <- c(epsilon = 0.1, alpha = 4, beta = 2)
pop <- dnorm(x.mids, mean=20, sd = 2) %>% {./sum(.)/dx}

@@

pop.inf <- steady.1D(func = ForagingMemoryModel, 
                   y = pop, parms = parameters, 
                   resource = resource, 
                   pop_lag = pop_lag, 
                   nspec=1, dx = dx)$y %>% {./sum(.)/dx}

@@

```{r setupEnvironment}
getSinePop <- function(Time, X, X.max, peak.max, peak.min, sd){
  pop <- outer(Time, X,
                     function(t,x) 
                       dnorm(x, mean = X.max/2 + (peak.max-peak.min) / 2 * sin( 2*(t * pi)/max(Time)),sd = sd))
  pop <- apply(pop, 1, function(x) x/sum(x)) %>% t
  pop / mean(diff(X))
}

dx <- 1
X.max <- 100
tau <- 100

X.edge <- seq(0,100, length=X.max/dx + 1)
X.mid <- seq(1,X.max,dx) - 1/2
Time <- 1:tau

pop.Year1 <- getSinePop(Time, X.mid, X.max = X.max, peak.max = 80, peak.min = 20, sd = 10)

fr1 <- dnorm(X.mid, mean = 80, sd = 10) %>% {./sum(.)/dx}
fr2 <- dnorm(X.mid, mean = 20, sd = 10) %>% {./sum(.)/dx}

fixed.Resource <- matrix((fr1 + fr2)/2, byrow = TRUE, 
                         nrow = nrow(pop.Year1), 
                         ncol = ncol(pop.Year1))

@@

```{r}
ForagingMemoryModel <- function(t, y, parms, resource, pop_lag){
  tran.1D(C = y, D = parms["epsilon"], 
          flux.up = 0, flux.down = 0, 
          v = parms["alpha"] * diff(resource)/parms["dx"] +  
            parms["beta0"] * diff(y)/parms["dx"] +
            parms["beta1"] * diff(pop_lag)/parms["dx"], 
          dx = parms["dx"])
}
```

```{r memoryOnly, fig.height = 5, fig.width = 10}
dx <- .5
parameters <- c(epsilon = 5, alpha = 0, 
                beta0 = 20, 
                beta1 = 5e2, 
                lambda = 0, dx = dx)

eval <- FALSE
if(eval){
  MemSim <- runManyYears(pop.Year1, Time, Resource = fixed.Resource, Parameters = parameters, n.years = 3)
  save(MemSim, file = "../results/MemSim.rda")
} else load("../results/MemSim.rda")

>>>>

---
title: "Modeling memory and taxis to analyze migrations"
author: "E. Gurarie, S. Potluri, B. Fagan, others"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
theme: cerulean
editor_options:
  chunk_output_type: console
---

@@

```{r, message=FALSE, warning=FALSE}
require(deSolve)
require(ReacTran)
require(magrittr)
```

@@

```{r}
ForagingMemoryModel <- function(t, pop, parms, pop_lag, resource, 
                                dx = attributes(pop)$dx){
  tran.1D(C = pop, D = parms["epsilon"], 
                   flux.up = 0, flux.down = 0, 
                   v = parms["alpha"] * diff(resource)/dx + 
                       parms["beta"] * diff(pop_lag)/dx, 
                   dx = dx)
}
```

@@

parameters <- c(epsilon = 0.1, alpha = 4, beta = 2)
pop <- dnorm(x.mids, mean=20, sd = 2) %>% {./sum(.)/dx} 
attributes(pop)$dx <- dx

@@

pop.inf <- steady.1D(func = ForagingMemoryModel, 
                   y = pop, parms = parameters, 
                   resource = resource, 
                   pop_lag = pop_lag, 
                   nspec=1, 
                   dx = attributes(pop)$dx)$y %>% {./sum(.)/attributes(pop)$dx}

@@

```{r setupEnvironment}
getSinePop <- function(tau, X.min = 0, X.max = 100, dx, 
                       peak.max, peak.min, sd){
  Time <- 1:tau

  X.edge <- seq(X.min, X.max, dx)
  X.mid <- (X.edge[-1] + X.edge[-length(X.edge)])/2
  pop <- outer(Time, X.mid,
                     function(t,x) 
                       dnorm(x, mean = X.max/2 + (peak.max-peak.min) / 2 * sin( 2*(t * pi)/max(Time)),sd = sd))
  pop <- apply(pop, 1, function(x) x/sum(x)/dx) %>% t
  list(pop = pop, X.mid = X.mid, time = Time, dx = dx, tau = tau)
}

world <- getSinePop(tau = 100, X.min = 0, X.max = 100, dx=1, 
                    peak.max = 80, peak.min = 20, sd = 10)

fr1 <- dnorm(world$X.mid, mean = 80, sd = 10) %>% {./sum(.)/dx}
fr2 <- dnorm(world$X.mid, mean = 20, sd = 10) %>% {./sum(.)/dx}
fixed.Resource <- matrix((fr1 + fr2)/2, byrow = TRUE, 
                         nrow = nrow(world$pop), 
                         ncol = ncol(world$pop))

world$resource <- fixed.Resource
```

@@

```{r memoryOnly, fig.height = 5, fig.width = 10}

parameters <- c(epsilon = 5, alpha = 0, beta0 = 20, beta1 = 5e2, lambda = 0)

eval <- FALSE
if(eval){
  MemSim <- runManyYears(World, Parameters = parameters, 
                         n.years = 3)
  save(MemSim, file = "../results/MemSim.rda")
} else load("../results/MemSim.rda")
